Preliminary Draft

### 2021-07-15 article/draft.tex
correct typos and add details

Title proposition 1: Big Tech is censoring me: using social media data to verify the platforms' regulation policies regarding ``misinformation''. 
%'Big Tech is censoring me': Using social media data to verify the platforms' regulation policies regarding ``misinformation''

@@

A number of recent studies  point towards the idea that ``Fake News'' or disinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense. \footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.''  }

@@

When an account is permanently suspended by Facebook, it disappears from the platform. That is,  the data can no longer be scrapped and it also disappears from the CrowdTangle API.\footnote{CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages.} Facebook publishes on monthly basis a {\it coordinated inauthentic behavior} report, where it informs how many personal accounts, pages or groups were deleted and to which {\it deceptive network} they may have belonged.\footnote{See the April 2021 report for an example: \href{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}} But as long as external persons do not have access to deleted accounts data, these reports cannot be verified by independent researchers or journalists.

Facebook can also apply a temporary suspension, and in this case the data can often be collected and analyzed. For example, \href{https://www.facebook.com/DonaldTrump/}{Donald Trump’s official Facebook page}  has been suspended following the Capitol attack on January 6, 2021.\footnote{See \href{https://www.facebook.com/zuck/posts/10112681480907401}{https://www.facebook.com/zuck/posts/10112681480907401}} Nevertheless the page’s data is still present in the CrowdTangle API. Thus, after manually adding this page to the CrowdTangle dashboard, we collected the $6$ $083$ posts it had published between January $1$, $2020$ and June $15$, $2021$ using the {\it posts} endpoint.\footnote{ (see the endpoint documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Posts}{https://github.com/CrowdTangle/API/wiki/Posts}.} We used the minet command line tool \cite{minet} to collect data.\footnote{The exact command can be found : \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_trump_data.sh}{here}.} We can verify on figure \ref{fig1_fb} that the {\it Donald J. Trump} page has not published any content since January $6$, 2021, and that this behavior is not consistent with the page’s previous activity: an average of $16$ posts were published each day on Facebook before the suspension.

@@

The Twitter account of the website lifesitenews.com has been suspended for at least two periods of time: from end of 2019 until fall 2020 for 308 days, then again since January 2021 for having violated Twitter Rules\footnote{See Lifesitenews's article discussing the reason for the suspension: \href{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}. Twitter rules can be found at: \href{https://help.twitter.com/en/rules-and-policies/twitter-rules}{https://help.twitter.com/en/rules-and-policies/twitter-rules}. }. In particular, this website has several failed fact-checks concerning the published articles, according to Iffy.news.\footnote{See \href{https://mediabiasfactcheck.com/life-site-news/}{https://mediabiasfactcheck.com/life-site-news/}}. We collected the activity (tweets, replies, quotes, retweets) on their Twitter account $@LifeSite$ via the Twitter API, using the historical search endpoint.\footnote{See the documentation: \href{https://developer.twitter.com/en/docs/twitter-api/tweets/search/api-reference/get-tweets-search-all}{https://developer.twitter.com/en/docs/twitter-api/tweets/search/api-reference/get-tweets-search-all}.} We then plotted the number of Tweets, Retweets, Quotes and Replies per day, as shown in panel $a$ of figure \ref{fig2}). The two periods of temporary suspension are clearly observed in the data as the user(s) of the account were not allowed to use the functionalities of the Twitter Platform.

@@

A third measure that mainstream social media platforms can apply is to prevent users from sharing specific types of content, instead of deleting the content after posting, or users accounts. We will show through examples how platforms prevent users from sharing urls coming from specific domain names.

@@

In this section, we provide one case study of how Facebook can block a URL as part of its policy enforcement tools. The Beauty of life (\href{https://thebl.com/}{thebl.com/}) is a US-based media company that shares pro-Trump views and conspiracy theories such as QAnon.\footnote{ See \href{https://en.wikipedia.org/wiki/The\_Epoch\_Times\#Removal\_of\_The\_BL\_(The\_Beauty\_of\_Life)\_from\_Facebook}{wikipedia article}.} Facebook has announced on December $20$, $2019$ that ``The BL is now banned from Facebook'' for coordinated inauthentic behavior\footnote{\href{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}.}, which includes using fake accounts that misrepresent one's identity or using methods to artificially boost the popularity of content. Coordinated inauthentic behavior is a distinct phenomenon from disinformation according to Facebook, as ``most of the content shared by coordinated manipulation campaigns isn’t probably false”.\footnote{See: \href{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}.}

Nevertheless, for the case of the Beauty of Life, both misinformation and coordinated inauthentic behavior are interlinked, according to the fact-checking organization Snopes which had reported about The BL’s activity to Facebook and other various public articles.\footnote{\href{ https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{ https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, \href{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, \href{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}} It took Facebook five months to complete its investigation of The BL before taking action\footnote{See \href{https://www.businessinsider.com/facebook-bans-beauty-of-life-for-inauthentic-behavior-2019-12}{https://www.businessinsider.com/facebook-bans-beauty-of-life-for-inauthentic-behavior-2019-12}.} and Snopes ended its fact-checking partnership with them in February 2020.\footnote{See \href{https://www.snopes.com/2019/02/01/snopes-fb-partnership-ends/}{https://www.snopes.com/2019/02/01/snopes-fb-partnership-ends/}.}

@@

To verify Facebook’s ban of The BL domain name, we first tested whether we could post a Facebook message containing a url from thebl.com. This turned out to be impossible. But such manual verification cannot inform us whether the ban applies indeed to all Facebook users and accounts (as we used only our own personal accounts), nor when it has started. To further investigate this policy, we collected data from Buzzsumo.\footnote{BuzzSumo is a commercial content database that tracks the volume of user interactions with internet content on Facebook, Twitter, and other social media platforms.} We used the ``/search/articles'' endpoint of the Buzzsumo API, to collect the engagement metrics $13$ $634$ articles crawled from the $thebl.com$ website between January $1$, $2019$ and June $15$, $2021$.\footnote{The command can be found in the following Github repository: \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_thebl\_data.py}{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_thebl\_data.py}.}

The number of Facebook reactions, shares and comments dropped to zero for TheBL’s articles published after December $1$, $2019$ (see figure \ref{fb_bl} top panel), indicating the start of the ban. We can note that although the ban was communicated in an article published\footnote{See \href{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us}{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us}.} on December $20$, $2019$, it seems to have actually started on December $1$, $2019$. The number of articles published per day was around $50$ until December $20$, $2019$. This number decreased drastically to reach around $5$ to $10$ articles published per day (see figure \ref{fb_bl} bottom panel). The communication around the ban may have discouraged The Beauty of Life to proceed with their activity. Using Buzzsumo data, a global ban over sharing a thebl.com URL link on Facebook was ascertained. It started on December $1$, $2019$, and was still enforced in June 2021.

>>>>

Title proposition 1: ``Big Tech is censoring me'': using social media data to verify the platforms' regulation policies regarding misinformation.

Title proposition 2: Illustrating Facebook, Twitter and YouTube regulation policies against misinformation with a few chosen examples

@@

A number of recent studies  point towards the idea that ``Fake News'' or disinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense. \footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.''}

@@

When an account is permanently suspended by Facebook, it disappears from the platform. That is,  the data can no longer be scrapped and it also disappears from the CrowdTangle API.\footnote{CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages.} Facebook publishes on monthly basis a {\it coordinated inauthentic behavior} report, where it informs how many personal accounts, pages or groups were deleted and to which {\it deceptive network} they may have belonged.\footnote{See the \href{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}{April 2021 report} as an example.} But as long as external persons do not have access to deleted accounts data, these reports cannot be verified by independent researchers or journalists.

Facebook can also apply a temporary suspension, and in this case the data can often be collected and analyzed. For example, \href{https://www.facebook.com/DonaldTrump/}{Donald Trump’s official Facebook page}  has been suspended following the Capitol attack on January 6, 2021.\footnote{See \href{https://www.facebook.com/zuck/posts/10112681480907401}{https://www.facebook.com/zuck/posts/10112681480907401}} Nevertheless the page’s data is still present in the CrowdTangle API. Thus, after manually adding this page to the CrowdTangle dashboard, we collected the $6$ $083$ posts it had published between January $1$, $2020$ and June $15$, $2021$ using the {\it posts} endpoint.\footnote{See the endpoint documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Posts}{https://github.com/CrowdTangle/API/wiki/Posts}.} We used the minet command line tool \cite{minet} to collect data.\footnote{The exact command can be found \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_trump_data.sh}{here}.} We can verify on figure \ref{fig1_fb} that the {\it Donald J. Trump} page has not published any content since January $6$, 2021, and that this behavior is not consistent with the page’s previous activity: an average of $16$ posts were published each day on Facebook before the suspension.

@@

The Twitter account $@LifeSite$ of the website lifesitenews.com has been suspended for at least two periods of time: from end of 2019 until fall 2020 for 308 days, then again since January 2021 for having violated Twitter Rules\footnote{See Lifesitenews's article discussing the reason for the suspension: \href{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}. Twitter rules can be found at: \href{https://help.twitter.com/en/rules-and-policies/twitter-rules}{https://help.twitter.com/en/rules-and-policies/twitter-rules}. }. In particular, this website has several failed fact-checks concerning the published articles, according to Iffy.news.\footnote{See \href{https://mediabiasfactcheck.com/life-site-news/}{https://mediabiasfactcheck.com/life-site-news/}}. We collected the activity (tweets, replies, quotes, retweets) on their Twitter account via the Twitter API, using the historical search endpoint.\footnote{See the documentation: \href{https://developer.twitter.com/en/docs/twitter-api/tweets/search/api-reference/get-tweets-search-all}{https://developer.twitter.com/en/docs/twitter-api/tweets/search/api-reference/get-tweets-search-all}.} We then plotted the number of Tweets, Retweets, Quotes and Replies per day, as shown in panel $a$ of figure \ref{fig2}). The two periods of temporary suspension are clearly observed in the data as the user(s) of the account were not allowed to use the functionalities of the Twitter Platform.

@@

Another measure that mainstream social media platforms can apply is to prevent users from sharing specific types of content, in this example URLs coming from a specific domain name.

@@

The Beauty of life (\href{https://thebl.com/}{thebl.com/}) is a US-based media company that shares pro-Trump views and conspiracy theories such as QAnon.\footnote{ See \href{https://en.wikipedia.org/wiki/The\_Epoch\_Times\#Removal\_of\_The\_BL\_(The\_Beauty\_of\_Life)\_from\_Facebook}{wikipedia article}.} Facebook has announced on December $20$, $2019$ that ``The BL is now banned from Facebook'' for coordinated inauthentic behavior\footnote{\href{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}.}, which includes using fake accounts that misrepresent one's identity or using methods to artificially boost the popularity of content. Coordinated inauthentic behavior is a distinct phenomenon from disinformation according to Facebook, as ``most of the content shared by coordinated manipulation campaigns isn’t probably false”.\footnote{See: \href{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}{https://about.fb.com/news/2019/10/inauthentic-behavior-policy-update/}.} Nevertheless, for the case of the Beauty of Life, both misinformation and coordinated inauthentic behavior were attested, according to the fact-checking organization Snopes which had reported about The BL’s activity to Facebook and other various public articles\footnote{\href{ https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, \href{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}{https://www.snopes.com/news/2019/11/12/bl-fake-profiles/}, \href{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}{https://www.snopes.com/news/2019/12/13/facebook-bl-cib/}}.

@@

To verify Facebook’s ban of The BL domain name, we first tested whether we could post a Facebook message containing a url from thebl.com. This turned out to be impossible. But such manual verification cannot inform us whether the ban applies indeed to all Facebook users and accounts (as we used only our own personal accounts), nor when it has started. To further investigate this policy, we collected data from the Buzzsumo API\footnote{BuzzSumo is a commercial content database that tracks the volume of user interactions with internet content on Facebook, Twitter, and other social media platforms.}. We used the ``/search/articles'' endpoint to collect the engagement metrics of the $13$ $634$ articles crawled from the $thebl.com$ website between January $1$, $2019$ and June $15$, $2021$.\footnote{The command can be found in the following Github repository: \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_thebl\_data.py}{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect\_facebook\_buzzsumo\_thebl\_data.py}.}

The number of Facebook reactions, shares and comments dropped to zero for TheBL’s articles published after December $1$, $2019$ (see figure \ref{fb_bl} top panel), indicating the start of the ban. We can note that although the ban was communicated in an article\footnote{See \href{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us}{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us}.} published on December $20$, $2019$, it seems to have actually started on December $1$, $2019$.

The communication around the ban appeared to have discouraged The Beauty of Life to proceed with their activity. Indeed the number of articles they published daily was around $50$ until December $20$, $2019$, when it decreased drastically to reach around $5$ to $10$ articles published per day (see figure \ref{fb_bl} bottom panel). Using Buzzsumo data, we ascertained that links from thebl.com were not shared on Facebook anymore. The ban started on December $1$, $2019$, and appeared to be still enforced in June 2021.
